Modeling Intervention Strategies for United States TB Control

### 2013-09-09 in_progress/presentations/paper/draft.tex
worked on adding appendix

% latex table generated in R 3.0.1 by xtable 1.7-1 package
% Sat Aug 24 18:08:52 2013
\begin{table}
\centering
\begin{tabular}{|r|cccc|}
  \hline
 & red20 & red35 & red50 & red65 \\ 
  \hline
2000 & ND & ND & ND & ND \\ 
  2025 & 67722.96 & 67527.15 & 67332.53 & 67138.90 \\ 
  2050 & 28759.00 & 28581.09 & 28404.13 & 28227.50 \\ 
  2075 & 20876.80 & 20709.13 & 20541.77 & 20374.49 \\ 
  2100 & 17967.39 & 17804.94 & 17642.64 & 17480.33 \\ 
   \hline
\end{tabular}
\caption{Cost Per Case Averted by Reducing Incoming LTBI by 20\%, 35\%, 50\%, and 65\% (in dollar per case)} 
\label{tab:cpcaRes}
\end{table}

\subsection{Agent Based Evaluation}
The Agent Based Model allowed the statistical properties of the system to be
analyzed and verified. In particular, they illustrated that the deterministic
Hill model provides a robust and consistent statistical measure of TB epidemic
behavior in the US conditions. We found that the distribution of incidence and
final population sizes were normal, with mean accurate to the deterministic
model and standard deviations given in table X, in the appendix.

\subsection{Sensitivity Analysis}
\section{Discussion}
These results confirm the hypothesis that curing incoming LTBI rates is a
necessary step towards elimination and indicate that it is a cost effective
option. 
\subsection{Future Work}
This work could be extended by examining different classes of interventions or
more accurately estimating intervention cost with the deterministic extended
Hill. Further work could also be done with the agent-based Hill Model, by using
it to examine the effect contact structure plays on US TB incidence levels or to
examine the effects drug-resistant TB will have on US TB dynamics. 
\end{document}

>>>>

% latex table generated in R 3.0.1 by xtable 1.7-1 package
% Sat Aug 24 18:08:52 2013
\begin{table}
\centering
\begin{tabular}{|r|cccc|}
  \hline
 & red20 & red35 & red50 & red65 \\ 
  \hline
2000 & ND & ND & ND & ND \\ 
  2025 & 67722.96 & 67527.15 & 67332.53 & 67138.90 \\ 
  2050 & 28759.00 & 28581.09 & 28404.13 & 28227.50 \\ 
  2075 & 20876.80 & 20709.13 & 20541.77 & 20374.49 \\ 
  2100 & 17967.39 & 17804.94 & 17642.64 & 17480.33 \\ 
   \hline
\end{tabular}
\caption{Cost Per Case Averted by Reducing Incoming LTBI by 20\%, 35\%, 50\%, and 65\% (in dollar per case)} 
\label{tab:cpcaRes}
\end{table}
Several other intervention strategies analyzed by the Hill Model were refined
with the Extended Hill, and economic properties about each of them were tracked.
The results for these interventions, which are less effective than curing
entering LTBI cases across the board, are given in the appendix. 
\subsection{Agent Based Evaluation}
The Agent Based Model allowed the statistical properties of the system to be
analyzed and verified. In particular, they illustrated that the deterministic
Hill model provides a robust and consistent statistical measure of TB epidemic
behavior in the US conditions. We found that the distribution of incidence and
final population sizes were normal, with mean accurate to the deterministic
model and standard deviations given in table X, in the appendix.

\subsection{Sensitivity Analysis}
\section{Discussion}
These results confirm the hypothesis that curing incoming LTBI rates is a
necessary step towards elimination and indicate that it is a cost effective
option. 
\subsection{Future Work}
This work could be extended by examining different classes of interventions or
more accurately estimating intervention cost with the deterministic extended
Hill. Further work could also be done with the agent-based Hill Model, by using
it to examine the effect contact structure plays on US TB incidence levels or to
examine the effects drug-resistant TB will have on US TB dynamics. 
\section{Appendix}
\subsection{Hill Constants}
Below, we detail some of the relevant constants in the Basic Hill Model. A full
listing of constants used in the original Hill Model can be found in CITATION!
\begin{figure}[h]
  \begin{center}
    \tiny{
    \begin{verbatim}
sigmaLBase  <- 0.057
fBase       <- 0.187
transBase   <- 1
incLTBIBase <- 1

#Treatment Effectiveness Data:
probHosp      <- .49 #Probability of hospitalization for active TB treatment
efficacyLTBI  <- .9  #LTBI treatement efficacy
adherenceLTBI <- .64 #LTBI treatment adherence
probLTBItreatsuccess <- efficacyLTBI*adherenceLTBI

#Cost Parameter Values: 
costtb   <- 2985   #TB treatment cost w/o hopsitalization <- Dylan supp. p.11-12
costhosp <- 25495  #TB treatment cost w/ hospitalization  <- Dylan supp. p.10
costLTBI <- 403.45 #LTBI treatment cost
Ct       <- costtb*(1-probHosp) + costhosp*probHosp #Cost of active TB treatment
Cl       <- costLTBI/probLTBItreatsuccess           #Cost of LTBI treatment

parms <- c(
mu0   = 1/78,      #Natural mortality rate USB per year
mu1   = 1/53,      #Natural mortality rate FB per year
ro    = 0.018,     #USB birth rate per year
alpha = 0.005,     #FB arrival rate per year
p     = 0.103,     #Fraction of new infections which are acute (fast
progressors)
vF    = 1.5,       #Progression rate of acute infection per year
l0    = 0.015,     #Prevalence of LTBI in the USB population in 2000
l1    = 0.211,     #Prevalence of LTBI in the FB population in 2000: 
r0    = 0.667,     #Fraction of cases due to reactivation in the USB population
r1    = 0.780,     #Fraction of cases due to reactivation in the FB population
vL0   = 0.0014,    #Progression rate for reactivation (chronic LTBI) in the USB
population per year
vL1   = 0.0010,    #Progression rate for reactivation (chronic LTBI) in the FB
population per year
q     = 0.708,     #Fraction of infections progressing to infectious disease
mud   = 0.115,     #Mortality rate due to TB per year
x     = 0.111,     #Fraction of re-infected chronic LTBI moving to acute
infection
ARI0  = 0.030/100, #Annual risk of infection for USB in 2000
beta  = 10.39,     #Effective contact rate per year
e0    = 0.965,     #Fraction of preferred contacts with own population for USB
e1    = 0.985,     #Fraction of preferred contacts with own population for FB
g     = 0.0047,    #Fraction of FB arrivals with LTBI who are fast progressors
phi0  = 1.114,     #Cumulative fraction self-cure and treatment of active
disease for both populations pre year RATES (USB)
phi1  = 1.167,     #Cumulative fraction self-cure and treatment of active
disease for both populations pre year RATES (FB)
sigmaF0 = 1.296,   #Cumulative fraction of treatment for acute infection for
both populations per year RATES (USB)
sigmaF1 = 1.301,   #Cumulative fraction of treatment for acute infection for
both populations per year RATES (FB)
sigmaLBase = sigmaLBase, #Treatment rate for chronic LTBI per year
fBase = fBase,      #Fraction of FB arrivals with LTBI

#2010 New Cases in Population i (millions)
#source: http://www.cdc.gov/mmwr/preview/mmwrhtml/mm5105a3.htm
newCases0 = .008714,  #US-born
newCases1 = .007554,  #Foreign-born

CtBase = Ct,
ClBase = Cl
)
    \end{verbatim}
  }
  \end{center}
  \caption{A list of relevant constants in the Basic \& Extended Hill Model}
  \label{fig:basicHillConstants}
\end{figure}<++>
\subsection{The Extended Hill Equations}
Below, we detail the full equations used in the extended hill model. These
equations extend those used in the standard Hill Model. In many cases, the
equations of the Hill model were split into their component parts, which could
then be tracked separately. 
\begin{figure}[h]
  \begin{center}
  \tiny{
    \begin{verbatim}
discV   <- 1/(1.03^t)  #amount costs, health states discount constant

#parameter values initialized for each time step
c01          <- (1-e0)*((1-e1)*N1)/((1-e0)*N0 + (1-e1)*N1)       #proportion of contacts made with FB individuals  (USB)
c00          <- 1 - c01                                          #proportion of contacts made with USB individuals (USB)
c10          <- (1-e1)*((1-e0)*N0)/((1-e0)*N0 + (1-e1)*N1)       #proportion of contacts made with USB individuals (FB)
c11          <- 1 - c10                                          #proportion of contacts made with FB individuals  (FB)
dLTBIEn      <- f*alpha*(N0+N1)                                  #FB arrivals with LTBI entering
dnatdeath0   <- mu0 * N0                                         #Natural deaths (USB)
dnatdeath1   <- mu1 * N1                                         #Natural deaths (FB)
dtbdeath0    <- mud * (I0 + J0)                                  #TB deaths (USB)
dtbdeath1    <- mud * (I1 + J1)                                  #TB deaths (FB)
dtbdeathD0   <- discV * dtbdeath0                                #TB deaths with discounting (USB)
dtbdeathD1   <- discV * dtbdeath1                                #TB deaths with discounting (FB)
dprogAcute0  <- vF*F0                                            #Acute LTBI progressions to Active TB disease (USB)
dprogAcute1  <- vF*F1                                            #Acute LTBI progressions to Active TB disease (FB)
dprogChron0  <- vL0*L0                                           #Chronic LTBI progressions to Active TB disease (USB)
dprogChron1  <- vL1*L1                                           #Chronic LTBI progressions to Active TB disease (FB)
dprogTotal0  <- dprogAcute0 + dprogChron0                        #Progression to Active TB (USB)
dprogTotal1  <- dprogAcute1 + dprogChron1                        #Progression to Active TB (FB)
dprogTotalD0 <- discV * dprogTotal0                              #Progression to Active TB with discounting (USB)
dprogTotalD1 <- discV * dprogTotal1                              #Progression to Active TB with discounting (FB)
lambda0      <- transmission*(beta*(c00*(I0/N0) + c01*(I1/N1)))  #Forces of Infection (USB)
lambda1      <- transmission*(beta*(c10*(I0/N0) + c11*(I1/N1)))  #Forces of Infection (FB)
dexogenous0	 <- x*p*lambda0*L0                                   #Exogenous re-infections of Chronic LTBI to Acute LTBI (USB)
dexogenous1  <- x*p*lambda1*L1                                   #Exogenous re-infections of Chronic LTBI to Acute LTBI (FB)
dInterventionCost <- discV * (iCnewCases*1e6*(dprogTotal0+dprogTotal1) + iCtotPop*1e6*(N0+N1) + iCLTBIEn*1e6*dLTBIEn*(1-incLTBI))

#Difference Equations (USB)
dS0 <- ro*(N0+N1) + sigmaF0*F0 + sigmaL*L0 + phi0*(I0+J0) - lambda0*S0 - mu0*S0
dF0 <- p*lambda0*S0 + dexogenous0 - (mu0 + vF + sigmaF0)*F0
dL0 <- (1-p)*lambda0*S0 - dexogenous0 - (mu0 + vL0 + sigmaL)*L0
dI0 <- q*(dprogAcute0 + dprogChron0) - (mu0 + mud + phi0)*I0
dJ0 <- (1-q)*(dprogAcute0 + dprogChron0) - (mu0 + mud + phi0)*J0

#Difference Equations (FB)
dS1 <- (1-incLTBI)*dLTBIEn+(1-f)*alpha*(N0+N1) + sigmaF1*F1 + sigmaL*L1 + phi1*(I1 + J1) - lambda1*S1 - mu1*S1
dF1 <- g*p*dLTBIEn*incLTBI + p*lambda1*S1 + dexogenous1 - (mu1 + vF + sigmaF1)*F1
dL1 <- (1-g*p)*dLTBIEn*incLTBI + (1-p)*lambda1*S1 - dexogenous1 - (mu1 + vL1 +sigmaL)*L1
dI1 <- q*(dprogAcute1 + dprogChron1) - (mu1 + mud + phi1)*I1
dJ1 <- (1-q)*(dprogAcute1 + dprogChron1) - (mu1 + mud + phi1)*J1

dN0 <- dS0 + dF0 + dL0 + dI0 + dJ0
dN1 <- dS1 + dF1 + dL1 + dI1 + dJ1

#Cost calculations
dcL0 <- discV * Cl * sigmaL  * 1e6 * L0                    #cost for Chronic LTBI cures      (USB)
dcF0 <- discV * Cl * sigmaF0 * 1e6 * F0                    #cost for Acute LTBI cures        (USB)
dcI0 <- discV * Ct * q*(dprogAcute0 + dprogChron0) * 1e6     #cost for Infectious TB cures     (USB)
dcJ0 <- discV * Ct * (1-q)*(dprogAcute0 + dprogChron0) * 1e6 #cost for Non-Infectious TB cures (USB)
dcI0dL0 <- discV * Ct * q*(dprogChron0) * 1e6     #cost for Infectious TB cures     (USB)
dcI0dF0 <- discV * Ct * q*(dprogChron0) * 1e6     #cost for Infectious TB cures     (USB)
dcJ0dL0 <- discV * Ct * (1-q)*(dprogAcute0) * 1e6 #cost for Non-Infectious TB cures (USB)
dcJ0dF0 <- discV * Ct * (1-q)*(dprogAcute0) * 1e6 #cost for Non-Infectious TB cures (USB)
dcL1 <- discV * Cl * sigmaL  * 1e6 * L1                    #cost for Chronic LTBI cures      (FB)
dcF1 <- discV * Cl * sigmaF1 * 1e6 * F1                    #cost for Acute LTBI cures        (FB)
dcI1 <- discV * Ct * q*(dprogAcute1 + dprogChron1) * 1e6     #cost for Infectious TB cures     (FB)
dcJ1 <- discV * Ct * (1-q)*(dprogAcute1 + dprogChron1) * 1e6 #cost for Non-Infectious TB cures (FB)
dcI1dL1 <- discV * Ct * q*(dprogChron1) * 1e6     #cost for Infectious TB cures     (USB)
dcI1dF1 <- discV * Ct * q*(dprogChron1) * 1e6     #cost for Infectious TB cures     (USB)
dcJ1dL1 <- discV * Ct * (1-q)*(dprogAcute1) * 1e6 #cost for Non-Infectious TB cures (USB)
dcJ1dF1 <- discV * Ct * (1-q)*(dprogAcute1) * 1e6 #cost for Non-Infectious TB cures (USB)
dcN0 <- dcF0 + dcL0 + dcI0 + dcJ0                            #Total cost for all treatments (USB)
dcN1 <- dcF1 + dcL1 + dcI1 + dcJ1                            #Total cost for all treatments (FB)
    \end{verbatim}
  }
  \end{center}
  \label{fig:extendedHillEqs}
  \caption{
    The equations defining the Extended Hill Model, shown in \texttt{R} syntax.
    In this model, compartments share the same names as they did in the Hill
    Model, \emph{i.e.} \texttt{S1} is the Foreign Born susceptible population. 
  }
\end{figure}
Here, a continuous discounting approximation is used to a discount rate of $3\%$
yearly, found via $discV$. In many of the differential equations, a small $c$ is
used to denote \emph{cost of}. 
\subsection{Estimations for Infectious Rates of FB and USB}
Colin to add. 
\subsection{Statistical Qualities of the Agent Based Model}
\subsection{Finer Intervention Analysis}
\subsection{Sensitivity Analysis}
\end{document}
